---
title: "Phase 1A LPA - STARS & RMARS"
author: "Jenny Terry"
date: "10/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)
library(magrittr)
```

First, read in the cleaned, anonymised data.

```{r import_data}
# read in anonymised data file
statanx_data <- readr::read_csv(here::here("data", "anon_data.csv"))

# check first few rows
head(statanx_data)
```

```{r composite_scores}
# create composite scores

# stars mean
statanx_data <- statanx_data %>% dplyr::mutate(stars = rowMeans(dplyr::select(., dplyr::starts_with("q4.1")), na.rm = TRUE))

# rmars mean
statanx_data <- statanx_data %>% dplyr::mutate(rmars = rowMeans(dplyr::select(., dplyr::starts_with("q4.3")), na.rm = TRUE))

```

```{r estimate_profiles}
# fit iterative models to estimate optimal no. of profiles
ph1a_lpa_mod <- statanx_data %>%
    dplyr::select(stars, rmars) %>% 
    scale() %>% # centre variables
    tidyLPA::single_imputation(method = "missForest") %>%
    tidyLPA::estimate_profiles(1:6) %>%
    tidyLPA::get_fit()
ph1a_lpa_mod
```

For model interpretation see: https://data-edu.github.io/tidyLPA/articles/Introduction_to_tidyLPA.html#getting-fit-statistics

```{r mod1}
# re-run model with retained no. of profiles
# equal variances and covariances fixed to zero (model 1, default)
ph1a_mod1 <- statanx_data %>%
    dplyr::select(stars, rmars) %>%
    scale() %>%
    tidyLPA::single_imputation(method = "missForest") %>%
    tidyLPA::estimate_profiles(6)
ph1a_mod1
```

```{r mod2}
# re-run model with retained no. of profiles
# varying variances and covariances fixed to zero (model 2)
ph1a_mod2 <- statanx_data %>%
    dplyr::select(stars, rmars) %>%
    scale() %>%
    tidyLPA::single_imputation(method = "missForest") %>%
    tidyLPA::estimate_profiles(6, variances = "varying", covariances = "zero")
ph1a_mod2
```

```{r mod3}
# re-run model with retained no. of profiles
# equal variances and equal covariances (model 3)
ph1a_mod3 <- statanx_data %>%
    dplyr::select(stars, rmars) %>%
    scale() %>%
    tidyLPA::single_imputation(method = "missForest") %>%
    tidyLPA::estimate_profiles(6, variances = "equal", covariances = "equal")
ph1a_mod3
```

NOTE: models 4 & 5 are only available in MPlus.

```{r mod6}
# re-run model with retained no. of profiles
# varying variances and varying covariances (model 6)
ph1a_mod6 <- statanx_data %>%
    dplyr::select(stars, rmars) %>%
    scale() %>%
    tidyLPA::single_imputation(method = "missForest") %>%
    tidyLPA::estimate_profiles(6, variances = "varying", covariances = "varying")
ph1a_mod6
```

```{r get_estimates}
tidyLPA::get_estimates(ph1a_mod1)
```

Create a profile plot according to best practices, focusing on the visualisation of classification uncertainty by showing:

1. Bars reflecting a confidence interval for the class centroids
2. Boxes reflecting the standard deviations within each class; a box encompasses +/- 64% of the
observations in a normal distribution
3. Raw data, whose transparency is weighted by the posterior class probability, such that each datapoint is most clearly visible for the class it is most likely to be a member of.

```{r extract_profiles}
# create object containing profile no. for each case
library(tidyLPA) # for some reason using the namespace before the get_data function doesn't work - package must be loaded in advance
profile <- get_data(ph1a_mod1)
get_data(ph1a_mod1)
```

```{r append_profile}
# add class variable to main dataset
statanx_data$ph1a_profile <- profile$Class
```

```{r profile_freq}
# no. of participants in each profile
dplyr::count(statanx_data, vars = ph1a_profile)
```

```{r poster_plot}
# rename variables
statanx_data %<>% dplyr::rename(`Statistics Anxiety` = stars,
                                `Maths Anxiety` = rmars)
# re-run model
ph1a_mod1 <- statanx_data %>%
    dplyr::select(`Statistics Anxiety`, `Maths Anxiety`) %>%
    scale() %>%
    tidyLPA::single_imputation(method = "missForest") %>%
    tidyLPA::estimate_profiles(6)

# plot retained model profiles
ph1a_lpa_plot <- tidyLPA::plot_profiles(ph1a_mod1, sd = T, ci = NULL) +
    ggplot2::labs(y = "Value (centred)", 
         title = "Values of statistics anxiety and maths anxiety in each retained profile") +
    ggplot2::theme_bw()
ph1a_lpa_plot
```
